---
title: "Class 11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Section 1: Introduction to the RCSB Protein Data Bank (PDB)

We first downloaded a CSV file from the website and we will import it.
```{r}
pdb_data <- read.csv("Data Export Summary.csv", row.names = 1)
```

Determine the percentage of structures solved by X-Ray and Electron Microscopy.
```{r}
totalstructures <- sum(pdb_data$Total)
xray_electron_sum <- (pdb_data[1,5] + pdb_data[3,5])
percentxrayelectron <- (xray_electron_sum / totalstructures) * 100
print(percentxrayelectron)
```


What percent of structures are protein?
```{r}
protein <- sum(pdb_data$Proteins)
percentprotein <- (protein / totalstructures) * 100
print(percentprotein)
```

##Section 2: Visualizing the HIV-1 protease structure
All of this section was done in VMD.

##Section3: Introduction to Bio3D in R
Load the bio3D package
```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hsg.pdb")
pdb
```

Let's convert the three letter amino acid sequence to a one letter sequence using the aa321 function.
```{r}
aa321(pdb$seqres)
```

```{r}
head(pdb$atom)
```

##Section 4: Atom selection

Select all C-alpha atoms (return their indices)
```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

Select residue 10
```{r}
inds <- atom.select(pdb, resno = 10)
inds
```

Note that the attributes of the returned ca.inds from atom.select() include both atom and
xyz components. These are numeric vectors that can be used as indices to access the
corresponding atom and xyz components of the input PDB structure object.

```{r}
head(pdb$atom[inds$atom, ])
```

Select protein only and write out a new pdb file (hsg_protein)
```{r}
hsg_protein <- atom.select(pdb, "protein", value = TRUE)
write.pdb(hsg_protein, file = "1hsg_protein.pdb")
```

Select ligand/drug and write out a new pdb file (hsg_ligand)
```{r}
hsg_ligand <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(hsg_ligand, file = "1hsg_ligand.pdb")
```

```{r}
library(bio3d.view)
```

Let's view the 3D structure
```{r}
view(pdb, "overview", col = "sse")
```

##Section 6: Working with multiple PDB files

Download some example PDB files
```{r}
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

Extract and align the chains we are interested in
```{r}
pdbs <- pdbaln(files, fit = TRUE)
```

Print to screen a summary of the 'pdbs' object
```{r}
pdbs
```

Access the first 5 rows, and 8 columns
```{r}
pdbs$ali[1:5, 1:8]

```
Associated residues numbers
```{r}

pdbs$resno[1:5, 1:8]

```

Calculate sequence identity
```{r}
seqidentity(pdbs)

```

Calculate RMSD
```{r}
rmsd(pdbs)
```


```{r}
# Calculate RMSD
rd <- rmsd(pdbs)
# Clustering
hc <- hclust(as.dist(rd))
grps <- cutree(hc, k=3)
# Plot results as dendrogram
hclustplot(hc, k=3)
```

Principle Component Analysis

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
# Plot our results summary (PCA score plot and scree-plot)
plot(pc.xray)
```

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")

```

