---
title: "Class 12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Structural Bioinformatics (Part 2)

## Section 1: In silico docking of drugs to HIV-1 protease

# 1.1 Obtaining and inspecting our input structure

```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```

Read the PDB file
```{r}
hiv <- read.pdb(file.name)
```

Get a quick summary of the pdb structure
```{r}
hiv
```

Q1: What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure? 
> HOH and MK1 are non-protein resid values in this structure.
> resid corresponds to the residues. 

# 1.2 Prepare initial protein and ligand input files

Make protein-only and ligand-only objects called prot and lig that you can then write out to a new PDB formal files.
```{r}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")
```

# 1.3 Using AutoDockTools to setup protein docking input

In ADT (AutoDocTools), load the protein using File > Read Molecule.
Select 1hsg_protein.pdb. Click Open

We opened out protein only PDB file in AutoDocTools and added hydrogens and atom-trpes needed for docking calculations. 

We will use AutoDoc Vina here at the UNIX command line!

# 1.4 Prepare the ligand

In ADT (AutoDocTools), load the ligand using File > Read Molecule.
Select 1hsg_ligand.pdb. Click Open

We opened out ligand only PDB file in AutoDocTools and added hydrogens and atom-trpes needed for docking calculations. 

# 1.5 Prepare a docking configuration file

We need to create an input file that defines the protein, ligand and the search parameters. We will create the input file in a text editor.

In the new text file, we will put the grid dimensions found in ADT.

## Section 2: Docking ligands into HIV-1 protease

We used a program called Autodock Vina ran it from the command line.

'"\Program Files (x86)\The Scripps Research Institute\Vina\vina.exe" --config config --log log.txt'

# 2.3 Inspecting your docking results

Process the all.pdbqt to a PDB formal file that can be loaded into VMD.
```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)
write.pdb(res, file = "results.pdb")
```

Now both the original '1hsg.pdb' and the new 'results.pdb' file can be loaded into VMD. 

To assess the results quantitatively we will calculate the RMSD between the docking results and the knock crystal structure.
```{r}
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori, res)
```

## Section 3: Exploring the conformational dynamics of proteins

# 3.1 Normal Mode Analysis (NMA)

```{r}
pdb <- read.pdb("1HEL")
modes <- nma(pdb)
m7 <- mktrj(modes, mode = 7, file = "nma_7.pdb")
plot(modes, sse = pdb)
```

