---
title: "Class06 Homework"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Q6. How would you generalize the original code above to work with any set of input protein structures?
First, we need to load the bio3d library to run the code

```{r}
library(bio3d)
```

## Original Code to be generalized:
```{r}
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

## First, we must change it to read easier with one vector "x"
```{r}
    # x <- read.pdb("x")
    # x.chainA <- trim.pdb(x, chain = "A", elety = "CA")
    # x.b <- x.chainA$atom$b
    # plotb3(x.b, sse = x.chainA, typ = "l", ylab = "Bfactor")
```

## Now we can write the function!

  This function analyzes protein drug interactions by reading in any protien PDB data and outputs 
  a plot for the specified protein showing the B-factor versus the residue of the protein. 
  The function has one argument "pdb.filename" which is the name of the PDB file to be analyzed. 

      The first statement reads the PDB file and restricts the original PDB file to only include the  
      atoms/chains we want to analyze. The argument MUST be in quotes.
      In this case, we are focusing our analysis on alpha carbons and creating an object "x.chainA"
      
      The second statement further specifies the rows of the trimmed PDB file to analyze. 
      In this case, we are focusing on the B-factor of the atoms in the trimmed PDB file we created in the 
      previous line and creating an object "x.b"
      
      The third statement produces the output of our function. 
      Here, we create a line plot of the protein's residue (from x.b) versus that atom's B-factor. 
      The plot also includes secondary structure analyses in the margins of the plot as black and grey lines.

```{r}
bfactor.plot <- function(pdb.filename) {

  x.chainA <- trim.pdb(read.pdb(pdb.filename), chain = "A", elety = "CA")

  x.b <- x.chainA$atom$b
  
  plotb3(x.b, sse = x.chainA, typ = "l", ylab = "Bfactor", main = pdb.filename)
}
```


# Does the function work?? Let's see!

```{r}
bfactor.plot("4AKE")
bfactor.plot("1AKE")
bfactor.plot("1E4Y")
```

