---
title: "Untitled"
author: "Gabriela Goldberg"
date: "4/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More on function writing!

First we will revisit our function from last class.

```{r}
source("http://tinyurl.com/rescale-R")
```

Test the **rescale()** function

```{r}
rescale(1:10)
```
```{r}
rescale( c(1:10, "string"))
```

Let's fix this with another line of code.

```{r}
rescale2( c(1:10, "string"))
```
Now the error will tell us what is wrong.

## Function practice!

Write a function to identify NA elements in two vectors.

Start with a simple example input where I know what the answer should be.

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

I am looking for the positions where it is TRUE in both vectors...

```{r}
is.na(x) & is.na(y)
```

Take the sum to find how many TRUEs

```{r}
sum(is.na(x) & is.na(y))
```

This is my working snippet of code that I can use as the body of my first function!

```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

TEST!!

```{r}
both_na(x, y)
```

```{r}
both_na( c(NA, NA, NA, 5, 6, 7, 9, NA), c(NA, 5, 6, 7, NA, NA, NA))
```

Check to see if the lengths of our inputs are equal

```{r}
x <- c(NA, NA, NA)
y <- c(a, NA, NA, NA, NA, NA)
length(x) != length(y)
```

Function **both_na2()** will tell us to make the vectors the same length.

```{r}
both_na2(x, y)
```

Function **both_na3()** makes the code better and gives us more information.

```{r}
x <- c(NA, NA, NA, 1, 2, 3)
y <- c(NA, 3, 4, 5, NA, NA)
both_na3(x, y)
```

## Your turn! 

Write a function **grade()** to determine an overall grade from a vector of student homework assignment scores dropping the lowest single alignment score.

Student 1
```{r}
s1 <- c(100, 100, 100, 100, 100, 100, 100, 100, 90)
```

Student 2
```{r}
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
```

Let's first find the minimum of each student's scores.

```{r}
min(s1)
min(s2)

```

We can then add up all the scores and subtract them from the total and divide them by the length of the vector minus 1.

```{r}
(sum(s1) - min(s1)) / (length(s1) - 1)
(sum(s2) - min(s2)) / (length(s2) - 1)
```

Let's write the function!

```{r}
grade <- function(x) {
  (sum(x) - min(x)) / (length(x) - 1)
}
```

Let's test it on student 1

```{r}
grade(s1)
```

Does it work on student 2?

```{r}
grade(s2)
```

Let's try to fix this with a condition statement...

```{r}
grade <- function(x) {
  (sum(x, na.rm = TRUE) - min(x, na.rm = TRUE)) / (length(x) - 1)
}
```

Does it work on student 2?

```{r}
grade(s2)
```

Great! Now let's try to do this on a bigger file!

```{r}
url <- "http://tinyurl.com/gradeinput"
```

```{r}
classgrades <- read.csv(url, row.names = 1)
```

This csv file has 20 rows (students) and 5 columns (grades).

```{r}
grade(classgrades[1,])
```

```{r}
finalgrades <- apply(classgrades, 1, grade)
```

```{r}
sort(finalgrades, decreasing = TRUE)
```

## Now let's do a real-world example

Find common genes in two data sets and return their associated data (from each data set)

Start with a simple version of the problem
```{r}
df1 <- data.frame(IDs=c("gene1", "gene2", "gene3"),
 exp=c(2,1,1),
 stringsAsFactors=FALSE)
df2 <- data.frame(IDs=c("gene2", "gene4", "gene3", "gene5"),
 exp=c(-2, NA, 1, 2),
 stringsAsFactors=FALSE)
```

simplify further to single vectors
```{r}
x <- df1$IDs
y <- df2$IDs
```

find the intersection of both vectors
```{r}
intersect(x, y)
```

That function doesn't give us the option to give us all the information along with the gene.

```{r}
x
y

x %in% y
x[x %in% y]
```

Let's bind these columns together

```{r}
gene_intersect <- function(x, y) {
  cbind( x[x %in% y], y[y %in% x])
}
```

